---
title: "Transforming the data format from the SMK-api to fit the R world"
author: "Max Odsbjerg Pedersen"
date: '2022-05-16'
output: html_document
---
# Preface
In this module we are going to look further into the data returned from the flourishing period and what we can do with it. If you're coming directly from the previous module you will already have loaded the neccessary packages and the relevant data. For good measure we will load the packages and the data as the first thing. If any of these steps need further explanation refer to the previous module.

## Loading packages
```{r, message=FALSE}
library(tidyverse)
library(jsonlite)
```

## Loading data


```{r}
flourishing_artworks <- tibble()

for (i in seq(0, 4000, 2000)) {
  response_url <- paste0("https://api.smk.dk/api/v1/art/search?keys=%2A&range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&offset=", i, "&rows=2000")
  response <- fromJSON(response_url)
  df <- response$items
  flourishing_artworks <- bind_rows(flourishing_artworks, df)
}
```




# Examining the data closely

```{r}
flourishing_artworks
```

We have 5082 rows, which means we have 5082 art works returned from the flourishing period. These 5082 art works have 74 columns containing various meta data about the art works. At first sight this looks very good. It is a data frame and a popular way to work in R is based on data frames. But at close inspection things are not as straight forward as they may seem. Let's say we want to examine how these 5082 art works disperses on artists. Shuffling through the columns one might expect to come across a column called "artist", "creator" or something like this. But this is not the case. In order to find information about the creator of the art works we need to shift our focus to the column "production".

```{r}
flourishing_artworks %>% 
  select(production)
```
So here we have data frames within rows. This makes things a bit harder since the data frame isn't rectangular. Let's see what's inside the first dataframe:

```{r}
# 1 row of the 11th column extracted
flourishing_artworks[1, 11][[1]]
```
So this is where the creator of the art work hides along alot of other information about the creator. In order to find out how the 5082 art works disperses on creators we will extract the creator value from each of the 5082 art work. But before we do so, let us examine another on of the rows:

```{r}
# 8th row of the 11th column extracted
flourishing_artworks[8, 11][[1]]
```
We see that this particular art work has not one but three creators. The idea of extracting the creator for all the 5082 art works seems to be imperfect as information will be lost. Nonetheless we will continueing this plan and only ectract the first row of the creator column for the 5082 art works. In order to achieve this we will use the `map()`-function in order to select the creator-column from each of the 5082 art works. Using `mutate()` we create a new column called "creator_1" 


```{r}
flourishing_artworks %>% 
  mutate(creator_1 = map(production, "creator"))  %>% 
  select(creator_1)
```
This has given us column that consists of character-vectors(<chr>). In the art works that only have one creator the length of the vector will be one, but in the case we saw before with three creators the lenth will be three. The next step is to extract the first element from these character-vectors thus extracting the name of the creator of the art works with just one creator noted and the first noted creator from art works with more than one. By using `mutate()` we change the column we created before using the `map_chr()`-function. This function is told to go through the "creator_1"-column and extract the first element of the character-vectors: 
```{r}
flourishing_artworks %>% 
  mutate(creator_1 = map(production, "creator"))  %>% 
  mutate(creator_1 = map_chr(creator_1, 1, .null = NA)) %>% 
  select(creator_1)
```

Since we have the creators of the art works now it is manageable to count the creators thus finding the dispersion of the art works on creators: 

```{r}
flourishing_artworks %>% 
  mutate(creator_1 = map(production, "creator"))  %>% 
  mutate(creator_1 = map_chr(creator_1, 1, .null = NA)) %>% 
  select(creator_1) %>% 
  count(creator_1, sort = TRUE)
```
But what about the titles of the art works. It turns out that the situation is similiar even though the column is actually called "titles". It still consists of dataframes within rows:
```{r}
#Row number 14 is "titles"
flourishing_artworks[14]
```

Again we can inspect the first one and see that the first column is called "title" and the row contains the title.
```{r}
#First row of the 14 th column ectracted
flourishing_artworks[1, 14][[1]]
```
How ever the situation is quite similiar på the creator-situation. We can have more than one title: 
```{r}
#First row of the 14 th column ectracted
flourishing_artworks[4, 14][[1]]
```

Just like before we can use the `map()`-function in conjunction with the `mutate()`-function to extract the titles:

```{r}
flourishing_artworks %>% 
  mutate(creator_1 = map(production, "creator"))  %>% 
  mutate(creator_1 = map_chr(creator_1, 1, .null = NA)) %>% 
  mutate(title = map(titles, "title")) %>% 
  mutate(title = map_chr(title, 1, .null = NA)) %>% 
  select(creator_1, title)
```

But what if we want a quick way to see the a piece of art work's information presented in a way that is more readable for people who, unlike you, don't tamper with APIs and raw data. A nice feature of this particular API is that i returns a link to each art work in the National Gellery's online platform: 

```{r}
flourishing_artworks %>% 
  select(frontend_url)
```

This column is all ready in the format we like. One observation pr. row. We do not need to process this column. So if we were to add the "frontend_url" to our dataframe with creator and titles it would be as simple as `select()` the column in the last line of code from before: 


```{r}
flourishing_artworks %>% 
  mutate(creator_1 = map(production, "creator"))  %>% 
  mutate(creator_1 = map_chr(creator_1, 1, .null = NA)) %>% 
  mutate(title = map(titles, "title")) %>% 
  mutate(title = map_chr(title, 1, .null = NA)) %>% 
  select(creator_1, title, frontend_url)
```

Information on when the production started and ended is stored in the column "production_date" and the situation is the same as when we extracted the creators and the titles:

```{r}
flourishing_artworks %>% 
  mutate(creator_1 = map(production, "creator"))  %>% 
  mutate(creator_1 = map_chr(creator_1, 1, .null = NA)) %>% 
  mutate(title = map(titles, "title")) %>% 
  mutate(title = map_chr(title, 1, .null = NA)) %>% 
  mutate(production_start = map(production_date, "start")) %>%
  mutate(production_start = map_chr(production_start, 1, .null = NA)) %>%
  mutate(production_end = map(production_date, "end")) %>%
  mutate(production_end = map_chr(production_end, 1, .null = NA)) %>%
  select(creator_1, title, frontend_url, on_display, production_start, production_end) -> flourishing_artworks_df
```

```{r}

library(lubridate)

flourishing_artworks_df %>% 
  mutate(date_end = as_date(production_end)) %>% 
  count(date_end)
```


```{r}
flourishing_artworks_df %>% 
  mutate(date_end = as_date(production_end)) %>% 
  count(date_end) %>% 
  ggplot(aes(x=date_end, y=n)) +
  geom_line() + 
  xlab("")
```
```{r}
flourishing_artworks_df %>% 
  mutate(date_start = as_date(production_start)) %>% 
  count(date_start) %>% 
  ggplot(aes(x=date_start, y=n)) +
  geom_line() + 
  xlab("")
```



```{r}
flourishing_artworks_df %>% 
  mutate(production_period = interval(production_start, production_end)) %>% 
  mutate(production_time = time_length(production_period, unit = "years")) %>% 
 count(production_time, sort = TRUE)
```



```{r}
flourishing_artworks_df %>% 
  mutate(production_period = interval(production_start, production_end)) %>% 
  mutate(production_time = time_length(production_period, unit = "years")) %>% 
  filter(near(production_time, 0.997260274))
```

So 1300 of the works have a production time of one year
