---
title: "Loading all art works from the flourishing period with the SMK API into R"
author: "Max Odsbjerg Pedersen"
date: '2022-05-02'
output: 
    html_document:
      df_print: paged
      toc: true
      toc_depth: 2
      toc_float: true
editor_options: 
  markdown: 
    wrap: 72
---
This particular lesson is written in R in the so called [R-markdown format](https://r4ds.had.co.nz/r-markdown.html). It is assumed that you have R and Rstudio installed. In this case you will be able to follow all steps by running the code in the grey boxes beneath. For further information on getting R and Rstudio see the [Prerequisites-section](https://r4ds.had.co.nz/introduction.html#prerequisites) of the book *R for Data Science*. 

This lesson is the second concrete example of how to interact with a specific API and after [interacting with the SMK API we are ready](https://github.com/maxodsbjerg/CHCAA_API_lesson/blob/main/API_examples/2_smk/1_smk_swagger/interacting_with_smk_range.md)


# Libraries

Before we embark on bringing in the data and doing a simple analysis we must load the relevant packages that adds numerous functionalisties to R - just as was done in the previous example. Here the relevant packages are:

```{r, message=FALSE}
library(tidyverse)
library(jsonlite)
```


# Loading the request url into R

From the previous lesson we have the request URL that returned 10 out of
5084 art works from the flourishing period. Let us import this into R:

```{r}
flourishing <- fromJSON("https://api.smk.dk/api/v1/art/search?keys=%2A&range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&offset=0&rows=10")
```

What we do in the code above is to take the data from the request URL
and put it into an element within R that we call "flourishing". In the
environment panel we see the new element: ![The element "flourishing" in
the
environment](pic/screenshots/loading_all_flourishing_works/0_environment.png)
By clicking on the name (not the blue button, but the actual name
"flourishing) we open a new tab, which let's us see what is inside this
element: ![Exploring element
"flourishing"](pic/screenshots/loading_all_flourishing_works/1_flouirishing_element.png)
When inspecting the "flourishing"-element we see the same information,
that we saw when we were looking at the response in the Swagger
interface of the SMK API, but here it is represented differently. To
recap: We have an offset of zero, we have ten rows meaning we got 10 arts
works in the response and there was found 5084. The next thing we notice
is "items", which consist of a data frame with 10 rows and 49 columns.
This is where all the information on the ten art works we got is stored.
We can pin point this particular part of the flourishing element for
inspection by first writing the name of the element and after a dollar
sign writing the name of the part that we want to inspect(in this case
"items"):

```{r}
colnames(flourishing$items)
```

So this is the interesting part and what we are after. But we still have
the problem that we only have ten art works out of 5084. Before we
venture into the data frame above(it poses challenges in it self), we
need to make it a data frame containing all the 5084 art works.

For us to do that we need to return our focus to the request link:

> <https://api.smk.dk/api/v1/art/search?keys=%2A&range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&offset=0&rows=10>

It looks like an incomprehensible blob of signs, but if we brake it down
it will begin to make sense:

The request URL basically consists of two parts:

Base:

> <https://api.smk.dk/api/v1/art/search>?

Request:

> keys=%2A&range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&offset=0&rows=10

The base URL remains the same no matter what, so let us focus on the
request part. This can be further broken down by looking at all the
ampersand(&)-signs. Every ampersand-sign indicates a new part of the
request. Let's break them apart one by one: \>keys=%2A&

So this is the search key, but what is "%2A"? This is because URLs can't
contain special characters since they are transmitted over the internet.
Therefore special characters needs to be encoded, so they can go into an
URL. In this case "\*" is translated into "%2A". So this particular part
of the request part is our asterisk search, which in most search engines
is equivalent to "give me everything". Lets look at the next part.

> range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&

This is the part were we state the period we are interested in. From the
previous lesson we recognise several part of it: the field we are
working with (production_dates_end), the start date (1775-01-01) and the
end date (1807-09-30). Besides that we see alot of percentage signs (%)
which is again due to URL encoding. Instead of explaing every signs
URL-encoding we can use the package "urltools"'s function `URLdecode()`
to decode it:

```{r}
library(urltools)

URLdecode("range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&")
```

What we get is the exact same thing that we inserted in the Swagger API
interface of the SMK API.

If we focus on the next part we see the following:

> offset=0&

This is were we define our offset, so were in the result the API should
begin to return. The next part of the request part of the URL is, where
we define how many rows that are returned: \>rows=10

So if we once again return to the entire request part of the URL:

> keys=%2A&range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&offset=0&rows=10

will in plain speak be:

> give us everything within the period starting from 1775-01-01 to
> 1807-09-30 and give us the first ten results

So if we return to our initial goal: getting all 5084 artworks instead
of just 10. Is the solution as simple as changing the rows part of the
request URL? Well lets try!

> <https://api.smk.dk/api/v1/art/search?keys=%2A&range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&offset=0&rows=5084>

When we try to put this request URL in the browser we are handed an
error message saying:

> {"error":"Too many rows, max is 2000"}

So this is a bit disappointing. Does this mean that we will never get
all 5084 art works from the period? Thankfully this is not the case. But
we need to tamper with both the rows and the offset part of the request
URL. Notice how the offset changes.

> <https://api.smk.dk/api/v1/art/search?keys=%2A&range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&offset=0&rows=2000>

> <https://api.smk.dk/api/v1/art/search?keys=%2A&range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&offset=2000&rows=2000>

> <https://api.smk.dk/api/v1/art/search?keys=%2A&range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&offset=4000&rows=2000>

The three URLs above are exactly the same except from the offset value.
In the first link we notice that it is still zero and that we have
changed rows to 2000 - the maximum we can get from one URL. This URL
will give us the first 2000 art works from the flouirshing period. In
the second URL the offset has been changed to 2000 with the rows still
being set to the maximum of 2000. This link thus takes up where the
first one left off. This is all so the logic in the third URL where the
offset has been set to 4000, since the two first urls returns the first
4000 art works from our period. These three URLs will return all of our
5084 art works from the period.

We can load these three links exactly the same way we did before:

```{r}
flourishing_0 <- fromJSON("https://api.smk.dk/api/v1/art/search?keys=%2A&range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&offset=0&rows=2000")

flourishing_2000 <- fromJSON("https://api.smk.dk/api/v1/art/search?keys=%2A&range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&offset=2000&rows=2000")

flourishing_4000 <- fromJSON("https://api.smk.dk/api/v1/art/search?keys=%2A&range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&offset=4000&rows=2000")

```

So now we have our 5084 dispersed on three elements. Remeber before how
the actual data was "hidden" inside *items* in the element and how we
pin pointed it with the dollar sig? We are going to do the same thing
and combine them with the `bind_rows()`-function and saving the result
to a data frame we call flourishing (thus overwriting our initial
element which was the response that only held 10 art works)

```{r}
flourishing <- bind_rows(flourishing_0$items, flourishing_2000$items, flourishing_4000$items)
```

Let us see our new data frame:

```{r}
flourishing
```

We've successfully extracted all the work from the flourishing period!
But a question might rise. What if we had chose a longer period
resulting i a larger body of art works being returned? This could create
a situation were we wanted to extract 20000 art works. This would have
us create 10 request URL each incrementing the offset by 2000. This
would be tedious work. Constructing URLs and other things in the hand is
fine, when dealing with smaller numbers like three, but you might want
to look into automating the task instead (both the creation of URL and
the extraction of the data frame). Lets extract our 5084 art works a
little smarter, which will be helpful the next time, where we might want
to extract many more art works:

We are going to use a for loop for this. If you need to freshen up your
knowledge on for loops see the [for loops and
range()](https://chcaa.au.dk/#/courses/Humanities_Programming/modules/Control_Flow#for%20loops%20and%20range())-section
in the [Humanities
Programming](https://chcaa.au.dk/#/courses/Humanities_Programming)-course.
Since the changing part of of URL is the offset value we use the
`seq()`-function to create a vector of our three offset-values.

```{r}
flourishing_artworks <- tibble()

for (i in seq(0, 4000, 2000)) {
  response_url <- paste0("https://api.smk.dk/api/v1/art/search?keys=%2A&range=%5Bproduction_dates_end%3A%7B1775-01-01T00%3A00%3A00Z%3B1807-09-30T00%3A00%3A00Z%7D&offset=", i, "&rows=2000")
  response <- fromJSON(response_url)
  df <- response$items
  flourishing_artworks <- bind_rows(flourishing_artworks, df)
}
```

In the code above we only need to change the `seq()`-function we we were
to construct 10 URLs returning 20000 art works. The `seq()`function
works like this: `seq(from, to, by)` So for the case with 20000 art
works we would insert into our for loop the following:

```{r}
seq(0, 20000, 2000)
```
